---
title: "Asymmetric_KaKs"
author: "qxs"
date: "August 27, 2017"
output:
  html_document: default
  pdf_document: default
---

# This script is used to analysis substitution rate difference between paleolog and non-paleolog in *Brassica rapa*
#### This script is a summary of all information
#### Aug 2017 by Xinshuai Qi
#### 


# 1) Define the function to merge all the information into one table for each group
#### This function will load all the input files and merge them into one table, included input files are:
#### mikeKaKs="brassica_codemloutput/codeml_outputchc_ath.txt"
#### hit="chc" # mike's term, need to be replaced
#### mikeindex="index/indices.PC106-127.scafSeq"
#### blast="blast_Brapa/PC106-127.blast.out"
```{r eval=FALSE}
getwd()
setwd("C:/Users/qxs/Desktop")
library(dplyr)
library(plyr)
library(grid)
library(ggplot2)
# #TN
# mikeKaKs="brassica_codemloutput/codeml_outputtrn_Ath.txt"
# hit="trn" # mike's term, need to be replaced
# mikeindex="index/indices.TN28-127.scafSeq"
# blast="blast_Brapa/TN28-127.blast.out"
# 
# #PC
# mikeKaKs="brassica_codemloutput/codeml_outputchc_ath.txt"
# hit="chc" # mike's term, need to be replaced
# mikeindex="index/indices.PC106-127.scafSeq"
# blast="blast_Brapa/PC106-127.blast.out"



###############################################
# define the function
# this function will load all the input files, including files such as:
# mikeKaKs="brassica_codemloutput/codeml_outputchc_ath.txt"
# hit="chc" # mike's term, need to be replaced
# mikeindex="index/indices.PC106-127.scafSeq"
# blast="blast_Brapa/PC106-127.blast.out"

eachgroup<-function(mikeKaKs,hit,mikeindex,blast){
  
# read Mike's data
TN.Ath<-read.table(mikeKaKs,sep="\t",header=T) 
colnames(TN.Ath)<-c("Queryath","Hit","Kaathvshit","Ksathvs","Ka.Ksathvshit")
# replace
TN.Ath$Hit<-gsub(hit,'',TN.Ath$Hit)   #you may need to replace this line
# only keep useful info
TN.Ath<-TN.Ath[,1:5]
head(TN.Ath)



#read Mike's gene code
TN.code<-read.table(mikeindex,sep="\t",header=F) 
colnames(TN.code)<-c("code","scaf")
head(TN.code)
# TN.code$V2<-gsub('>','',TN.code$V2) 
# TN.code$V2<-gsub(' ','  ',TN.code$V2) 
# #split V2
# library(stringr)
# TN.code$V2<-str_split_fixed(TN.code$V2, "  ", 4)
# ncol(TN.code)
# TN.code<-subset(TN.code, select=c("V1","V2.1"))



# read blast out
TN.blast<-read.table(blast,sep="\t",header=F) 
TN.blast$V2<-gsub('.*(Bra\\d*)\\|.*', '\\1',TN.blast$V2) 
TN.blast<-TN.blast[,c(1:2)]
colnames(TN.blast)<-c('scaf','Bra')
#each scaffold seq may has multiple match in the B. rapa genome.
# so I will only keep the first matches.
TN.blast<-TN.blast[!duplicated(TN.blast$scaf),]
head(TN.blast)


#paleolog annotation
paleolog<-read.table("paleolog.txt")
colnames(paleolog)<-c("Bra","paleolog")
head(paleolog)



#merge all data into one
All<-merge(TN.code,TN.Ath,by.x="code",by.y = "Hit",all.y=F)
All<-merge(All,TN.blast,by.x="scaf",by.y = "scaf",all.y=FALSE,all.x=FALSE)
All<-merge(All,paleolog,by.x="Bra",by.y = "Bra",all.y=FALSE,all.x=FALSE)
head(All)

All<-unique(All)
return(All)
}
###############################################
```

# 2) Using the defined function generate a merged table for each tested group.
```{r eval=FALSE}
##### vs Ath
#TN
Ath_TN<-eachgroup( "brassica_codemloutput/codeml_outputtrn_ath.txt","trn",
                  "index/indices.TN28-127.scafSeq","blast_Brapa/TN28-127.blast.out");head(Ath_TN)
#PC pak choi
Ath_PC<-eachgroup( "brassica_codemloutput/codeml_outputpak_ath.txt","pak",    
                   "index/indices.PC106-127.scafSeq","blast_Brapa/PC106-127.blast.out");head(Ath_PC)

#CC
Ath_CC<-eachgroup( "brassica_codemloutput/codeml_outputchc_ath.txt","chc",    
                   "index/indices.CC68-127.scafSeq","blast_Brapa/CC68-127.blast.out");head(Ath_PC)

#IS sarson
Ath_IS<-eachgroup( "brassica_codemloutput/codeml_outputsar_ath.txt","sar",    
                   "index/indices.IS136-127.scafSeq","blast_Brapa/IS136-127.blast.out");head(Ath_PC)

#RP rapini
Ath_RP<-eachgroup( "brassica_codemloutput/codeml_outputrap_ath.txt","rap",    
                   "index/indices.RP153-127.scafSeq","blast_Brapa/RP153-127.blast.out");head(Ath_PC)




##### vs Aar
#TN
Aar_TN<-eachgroup( "brassica_codemloutput/codeml_outputtrn_aar.txt","trn",
                   "index/indices.TN28-127.scafSeq","blast_Brapa/TN28-127.blast.out");head(Aar_TN)
#PC pak choi
Aar_PC<-eachgroup( "brassica_codemloutput/codeml_outputpak_aar.txt","pak",    
                   "index/indices.PC106-127.scafSeq","blast_Brapa/PC106-127.blast.out");head(Aar_PC)

#CC
Aar_CC<-eachgroup( "brassica_codemloutput/codeml_outputchc_aar.txt","chc",    
                   "index/indices.CC68-127.scafSeq","blast_Brapa/CC68-127.blast.out");head(Aar_PC)

#IS sarson
Aar_IS<-eachgroup( "brassica_codemloutput/codeml_outputsar_aar.txt","sar",    
                   "index/indices.IS136-127.scafSeq","blast_Brapa/IS136-127.blast.out");head(Aar_PC)

#RP rapini
Aar_RP<-eachgroup( "brassica_codemloutput/codeml_outputrap_aar.txt","rap",    
                   "index/indices.RP153-127.scafSeq","blast_Brapa/RP153-127.blast.out");head(Aar_PC)

save(Ath_TN, file="C:/Users/qxs/Desktop/RData/Ath_TN.RData")
save(Ath_RP, file="C:/Users/qxs/Desktop/RData/Ath_RP.RData")
save(Ath_IS, file="C:/Users/qxs/Desktop/RData/Ath_IS.RData")
save(Ath_PC, file="C:/Users/qxs/Desktop/RData/Ath_PC.RData")
save(Ath_CC, file="C:/Users/qxs/Desktop/RData/Ath_CC.RData")

save(Aar_TN, file="C:/Users/qxs/Desktop/RData/Aar_TN.RData")
save(Aar_RP, file="C:/Users/qxs/Desktop/RData/Aar_RP.RData")
save(Aar_IS, file="C:/Users/qxs/Desktop/RData/Aar_IS.RData")
save(Aar_PC, file="C:/Users/qxs/Desktop/RData/Aar_PC.RData")
save(Aar_CC, file="C:/Users/qxs/Desktop/RData/Aar_CC.RData")


```

# The head of each table are like this:
```{r}
load(file="C:/Users/qxs/Desktop/RData/Ath_TN.RData")
head(Ath_TN)

```
 
 

# 3) Compare the mean(Ka/Ks) value between paleolog and non-paleolog in each *Brassica rapa* groups
```{r}
library(plyr)
library(ggplot2)
load(file="C:/Users/qxs/Desktop/RData/Ath_TN.RData")
load(file="C:/Users/qxs/Desktop/RData/Ath_RP.RData")
load(file="C:/Users/qxs/Desktop/RData/Ath_IS.RData")
load(file="C:/Users/qxs/Desktop/RData/Ath_PC.RData")
load(file="C:/Users/qxs/Desktop/RData/Ath_CC.RData")

load(file="C:/Users/qxs/Desktop/RData/Aar_TN.RData")
load(file="C:/Users/qxs/Desktop/RData/Aar_RP.RData")
load(file="C:/Users/qxs/Desktop/RData/Aar_IS.RData")
load(file="C:/Users/qxs/Desktop/RData/Aar_PC.RData")
load(file="C:/Users/qxs/Desktop/RData/Aar_CC.RData")

mu <- ddply(Ath_TN, "paleolog", summarise, grp.mean=mean(Ka.Ksathvshit,na.rm = T)) ;mu
mu <- ddply(Ath_RP, "paleolog", summarise, grp.mean=mean(Ka.Ksathvshit,na.rm = T)) ;mu 
mu <- ddply(Ath_IS, "paleolog", summarise, grp.mean=mean(Ka.Ksathvshit,na.rm = T)) ;mu 
mu <- ddply(Ath_PC, "paleolog", summarise, grp.mean=mean(Ka.Ksathvshit,na.rm = T)) ;mu
mu <- ddply(Ath_CC, "paleolog", summarise, grp.mean=mean(Ka.Ksathvshit,na.rm = T)) ;mu

mu <- ddply(Aar_TN, "paleolog", summarise, grp.mean=mean(Ka.Ksathvshit,na.rm = T)) ;mu
mu <- ddply(Aar_RP, "paleolog", summarise, grp.mean=mean(Ka.Ksathvshit,na.rm = T)) ;mu
mu <- ddply(Aar_IS, "paleolog", summarise, grp.mean=mean(Ka.Ksathvshit,na.rm = T)) ;mu
mu <- ddply(Aar_PC, "paleolog", summarise, grp.mean=mean(Ka.Ksathvshit,na.rm = T)) ;mu
mu <- ddply(Aar_CC, "paleolog", summarise, grp.mean=mean(Ka.Ksathvshit,na.rm = T)) ;mu
```

# 4) Generate plot for each *B. rapa* groups
```{r fig11, fig.height = 10, fig.width = 10,fig.align = "center"}
##########  Now I have table for each dataset
library(ggplot2)
load(file="C:/Users/qxs/Desktop/RData/Ath_TN.RData")
load(file="C:/Users/qxs/Desktop/RData/Ath_RP.RData")
load(file="C:/Users/qxs/Desktop/RData/Ath_IS.RData")
load(file="C:/Users/qxs/Desktop/RData/Ath_PC.RData")
load(file="C:/Users/qxs/Desktop/RData/Ath_CC.RData")

load(file="C:/Users/qxs/Desktop/RData/Aar_TN.RData")
load(file="C:/Users/qxs/Desktop/RData/Aar_RP.RData")
load(file="C:/Users/qxs/Desktop/RData/Aar_IS.RData")
load(file="C:/Users/qxs/Desktop/RData/Aar_PC.RData")
load(file="C:/Users/qxs/Desktop/RData/Aar_CC.RData")

completeFun <- function(data, desiredCols) {
  completeVec <- complete.cases(data[, desiredCols])
  return(data[completeVec, ])
}


# WGT_f<-filter(Ath_TN,paleolog=="WGT-paleolog")
# mean(WGT_f$Ka.Ksathvshit,na.rm = T)
# non_f<-filter(Ath_TN,paleolog=="Non-paleolog")
# mean(non_f$Ka.Ksathvshit,na.rm = T)
# 
# library(plyr)

kaksplot<-function(Ath_TN)
  {
Ath_TN <- completeFun(Ath_TN, "paleolog")
mu <- ddply(Ath_TN, "paleolog", summarise, grp.mean=mean(Ka.Ksathvshit,na.rm = T)) #calculate mean value 
#head(mu)
plot_Ath_TN<-ggplot(Ath_TN,aes(x = Ka.Ksathvshit,color=paleolog,fill=paleolog))+
  geom_density(alpha=.3)+xlim(0,1)+
  ggtitle("Ath_vs_??") + theme(plot.title = element_text(lineheight=.8, face="bold")) +
  geom_vline(data=mu, aes(xintercept=grp.mean,color=paleolog),linetype="dashed")+
  annotate("text", x = 0.5, y = 4,   label = paste("mean WGT:",round(mu$grp.mean[2],3)),color="#00BFC4")+
  annotate("text", x = 0.5, y = 3, label = paste("mean Non:",round(mu$grp.mean[1],3)),color="#F8766D")+
  labs(x="Ka/Ks")
return(plot_Ath_TN)
 }

plot_Ath_TN<-kaksplot(Ath_TN)+ggtitle("Ath_vs_TN");plot_Ath_TN
plot_Ath_RP<-kaksplot(Ath_RP)+ggtitle("Ath_vs_RP");plot_Ath_RP
plot_Ath_IS<-kaksplot(Ath_IS)+ggtitle("Ath_vs_IS");plot_Ath_IS
plot_Ath_PC<-kaksplot(Ath_PC)+ggtitle("Ath_vs_PC");plot_Ath_PC
plot_Ath_CC<-kaksplot(Ath_CC)+ggtitle("Ath_vs_CC");plot_Ath_CC

plot_Aar_TN<-kaksplot(Aar_TN)+ggtitle("Aar_vs_TN");plot_Aar_TN
plot_Aar_RP<-kaksplot(Aar_RP)+ggtitle("Aar_vs_RP");plot_Aar_RP
plot_Aar_IS<-kaksplot(Aar_IS)+ggtitle("Aar_vs_IS");plot_Aar_IS
plot_Aar_PC<-kaksplot(Aar_PC)+ggtitle("Aar_vs_PC");plot_Aar_PC
plot_Aar_CC<-kaksplot(Aar_CC)+ggtitle("Aar_vs_CC");plot_Aar_CC
 

# A summarize plot of all the plots  
 
#############################################################################
# Define Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)
  
  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)
  
  numPlots = length(plots)
  
  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                     ncol = cols, nrow = ceiling(numPlots/cols))
  }
  
  if (numPlots==1) {
    print(plots[[1]])
    
  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    
    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
#############################################################################
 

# A summarize plot of all the plots
 
library(ggplot2)
multiplot(plot_Ath_TN, plot_Ath_RP,plot_Ath_IS,plot_Ath_PC, plot_Ath_CC,
          plot_Aar_TN, plot_Aar_RP,plot_Aar_IS,plot_Aar_PC, plot_Aar_CC,cols=2)
```

**Summary**
According to the above plot, we didn't observe higher Ka/Ks/mean(Ka/Ks) in paleologs than in non-paleologs across the tested groups.

**Discussion**
* The plot of TN-Ath group is an outlier, Non-paleologs have much higher mean Ka/Ks than the WGT-paleolog group. 
  * It's not clear why. But at least the maximum Ka



# Discussion and my other thoughts

#### **Total number of observed Ka/Ks comparison**
* TN-Ath	9262
* RP-Ath	11904
* IS-Ath	11408
* PC-Ath	12428
* CC-Ath	11280

* TN-Aar	9076
* RP-Aar	9769
* IS-Aar	9934
* PC-Aar	10727
* CC-Aar	9663

#### **Number of WGT-paleolog, non-paleolog, and their ratio**
```{r}
Num_paleolog<-function(x){
  L1<-length(which(x$paleolog == "WGT-paleolog"))
  L2<-length(which(x$paleolog == "Non-paleolog"))
  out<-paste("Num_WGT:",L1, "Num_Non:", L2,"ratio:", L1/L2)
  return(out)
}


Num_paleolog(Ath_TN)
Num_paleolog(Ath_RP)
Num_paleolog(Ath_IS)
Num_paleolog(Ath_PC)
Num_paleolog(Ath_CC)

Num_paleolog(Aar_TN)
Num_paleolog(Aar_RP)
Num_paleolog(Aar_IS)
Num_paleolog(Aar_PC)
Num_paleolog(Aar_CC)
```

* "Num_WGT: 9041 Num_Non: 221 ratio: 40.9095022624434"
* "Num_WGT: 10051 Num_Non: 260 ratio: 38.6576923076923"
* "Num_WGT: 9705 Num_Non: 245 ratio: 39.6122448979592"
* "Num_WGT: 10534 Num_Non: 274 ratio: 38.4452554744526"
* "Num_WGT: 9551 Num_Non: 234 ratio: 40.8162393162393"

* "Num_WGT: 7695 Num_Non: 178 ratio: 43.2303370786517"
* "Num_WGT: 8218 Num_Non: 209 ratio: 39.3205741626794"
* "Num_WGT: 8441 Num_Non: 200 ratio: 42.205"
* "Num_WGT: 9085 Num_Non: 236 ratio: 38.4957627118644"
* "Num_WGT: 8153 Num_Non: 199 ratio: 40.9698492462312"

**Discussion**
The ratio of WGT vs Non is about 40:1 
which is way higher than what we estimated 68:13
which suggest for the de novo assembly, there is a bias towards the WGT-paleolog 
because these genes have more copies, de novo assembly can easily capture these genes.
On the other hand, there is a reverse bias towards the non-paleologs.
this bias may leads to a higher assembly error rate in the non-paleologs. 

#### **Compare the mean(Ka), mean(Ks)** 
```{r}
library(plyr)
library(ggplot2)
load(file="C:/Users/qxs/Desktop/RData/Ath_TN.RData")
load(file="C:/Users/qxs/Desktop/RData/Ath_RP.RData")
load(file="C:/Users/qxs/Desktop/RData/Ath_IS.RData")
load(file="C:/Users/qxs/Desktop/RData/Ath_PC.RData")
load(file="C:/Users/qxs/Desktop/RData/Ath_CC.RData")

load(file="C:/Users/qxs/Desktop/RData/Aar_TN.RData")
load(file="C:/Users/qxs/Desktop/RData/Aar_RP.RData")
load(file="C:/Users/qxs/Desktop/RData/Aar_IS.RData")
load(file="C:/Users/qxs/Desktop/RData/Aar_PC.RData")
load(file="C:/Users/qxs/Desktop/RData/Aar_CC.RData")

# compare the Ks value
mu <- ddply(Ath_TN, "paleolog", summarise, grp.mean=mean(Ksathvs,na.rm = T)) ;mu
mu <- ddply(Ath_RP, "paleolog", summarise, grp.mean=mean(Ksathvs,na.rm = T)) ;mu 
mu <- ddply(Ath_IS, "paleolog", summarise, grp.mean=mean(Ksathvs,na.rm = T)) ;mu 
mu <- ddply(Ath_PC, "paleolog", summarise, grp.mean=mean(Ksathvs,na.rm = T)) ;mu
mu <- ddply(Ath_CC, "paleolog", summarise, grp.mean=mean(Ksathvs,na.rm = T)) ;mu

mu <- ddply(Aar_TN, "paleolog", summarise, grp.mean=mean(Ksathvs,na.rm = T)) ;mu
mu <- ddply(Aar_RP, "paleolog", summarise, grp.mean=mean(Ksathvs,na.rm = T)) ;mu
mu <- ddply(Aar_IS, "paleolog", summarise, grp.mean=mean(Ksathvs,na.rm = T)) ;mu
mu <- ddply(Aar_PC, "paleolog", summarise, grp.mean=mean(Ksathvs,na.rm = T)) ;mu
mu <- ddply(Aar_CC, "paleolog", summarise, grp.mean=mean(Ksathvs,na.rm = T)) ;mu

## compare the Ka value 
mu <- ddply(Ath_TN, "paleolog", summarise, grp.mean=mean(Kaathvshit,na.rm = T)) ;mu
mu <- ddply(Ath_RP, "paleolog", summarise, grp.mean=mean(Kaathvshit,na.rm = T)) ;mu 
mu <- ddply(Ath_IS, "paleolog", summarise, grp.mean=mean(Kaathvshit,na.rm = T)) ;mu 
mu <- ddply(Ath_PC, "paleolog", summarise, grp.mean=mean(Kaathvshit,na.rm = T)) ;mu
mu <- ddply(Ath_CC, "paleolog", summarise, grp.mean=mean(Kaathvshit,na.rm = T)) ;mu

mu <- ddply(Aar_TN, "paleolog", summarise, grp.mean=mean(Kaathvshit,na.rm = T)) ;mu
mu <- ddply(Aar_RP, "paleolog", summarise, grp.mean=mean(Kaathvshit,na.rm = T)) ;mu
mu <- ddply(Aar_IS, "paleolog", summarise, grp.mean=mean(Kaathvshit,na.rm = T)) ;mu
mu <- ddply(Aar_PC, "paleolog", summarise, grp.mean=mean(Kaathvshit,na.rm = T)) ;mu
mu <- ddply(Aar_CC, "paleolog", summarise, grp.mean=mean(Kaathvshit,na.rm = T)) ;mu


```
**Discussion**
In Non-paleologs have both higher mean Ka and higher mean Ks than WGT-paleologs acrossing the 10 datasets.





